An equivalence of compositional frameworks

### 2019-05-03 EquivCompFrameworks_20180911kc.tex
cleaned up preamble a bit and commented some stuff out after the equivalence theorem

We can also define the part of the double equivalence $\mathbb{G} \colon F\mathbb{C}\textnormal{sp} \to {_L \mathbb{C} \textnormal{sp}(\mathsf{X})}$ which goes in the other direction: again, the object component of this double functor will be $\mathbb{G}_0 = \id_{\mathsf{A}}$.

@@

On the other hand, given a horizontal 1-cell $N$ in $F\mathbb{C}\textnormal{sp}$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$c_2$};
\node (D) at (1,-0.5) {$x \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I$} (B)
(C) edge node[above]{$O$} (B);
\end{tikzpicture}
\]
the horizontal 1-cell $\mathbb{G}(N)$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(c_1)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(c_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$!_c L(I)$} (B)
(C) edge node[above]{$!_c L(O)$} (B);
\end{tikzpicture}
\]
and then the horizontal 1-cell $\mathbb{E} \mathbb{G}(N)$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1.5,0) {$R(x)$};
\node (C) at (3,0) {$c_2$};
\node (D) at (1.5,-0.5) {$x \in F(R(x))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$R(!_c L(I))\eta_{c_1}$} (B)
(C) edge node[above]{$R(!_c L(O))\eta_{c_2}$} (B);
\end{tikzpicture}
\]
Then we have a 2-isomorphism $\epsilon_N \colon \mathbb{E} \mathbb{G} (N) \xrightarrow{\sim} N$ in $F\mathbb{C}\textnormal{sp}$ given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (D) at (1.5,0.5) {$x \in F(R(x))$};
\node (D') at (1.5,-1.5) {$x \in F(c)$};
\node (A) at (0,0) {$c_1$};
\node (B) at (1.5,0) {$R(x)$};
\node (C) at (3,0) {$c_2$};
\node (A') at (0,-1) {$c_1$};
\node (B') at (1.5,-1) {$c$};
\node (C') at (3,-1) {$c_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$R(!_c L(I))\eta_{c_1}$} (B)
(C) edge node[above]{$R(!_c L(O))\eta_{c_2}$} (B)
(A') edge node[above]{$I$} (B')
(C') edge node[above]{$O$} (B')
(A) edge node [left]{$1$} (A')
(B) edge node [left]{$e$} (B')
(C) edge node [left]{$1$} (C');
\end{tikzpicture}
\]
$$\iota \colon F(e)(x) \to x$$
where 
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1,0) {$R(L(c_1))$};
\node (C) at (2.5,0) {$R(L(c))$};
\node (D) at (3.5,0) {$R(x)$};
\node (E) at (4.5,0) {$c$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$\eta_{c_1}$} (B)
(B) edge node[above]{$R(L(I))$} (C)
(C) edge node[above]{$R(!_c)$} (D)
(D) edge node[above]{$e$} (E)
(A) edge[bend right] node [above]{$I$} (E);
\end{tikzpicture}
\]
commutes.

\begin{comment}
\section{Applications}\label{Applications}
In this section we present several examples each of which may be realized in the context of decorated cospans or in the context of structured cospans. The first example regarding graphs was mentioned in the introduction and used as a reoccurring theme throughout the paper. The next three examples which take on more of an applied flavor, consists of electrical circuits, Markov processes and Petri nets. Each of these has been studied extensively by Baez, Fong, Master and Pollard by way of `black-boxing' \cite{BCR,BF,BFP,BM,BP}. Black-boxing is a way of interpreting the behavior of an open system, that is, a system with prescribed inputs and outputs such as the terminals of an electrical circuit, by observing the activity at the inputs and the outputs. The semantics of the activity at an open system's inputs and outputs is typically described in a category such as $\mathsf{LinRel}$ of finite dimensonal vector spaces and linear relations. Thus, in each case, black-boxing results in functors such as: $$\blacksquare_1 \colon \mathsf{Circ} \to \mathsf{LinRel}$$ $$\blacksquare_2 \colon \mathsf{Mark} \to \mathsf{LinRel}$$ $$\blacksquare_3 \colon \mathsf{Petri} \to \mathsf{LinRel}.$$ Each of these black-boxing functors also possess other convenient properties such as being symmetric monoidal. The first two of these were first done using Fong's theory of decorated cospans and then extended by the first two authors using structured cospans. The last two of these were also extended by being realized as double functors between double categories \cite{BC,BM} and is discussed in Chapter \ref{Chapter6}.
\subsection{Graphs}
As a first example that was also mentioned in the introduction, let $L \colon \mathsf{Set} \to \mathsf{Graph}$ be the functor that assigns to a set $N$ the \emph{discrete graph} on $N$ which is the edgeless graph $L(N)$ with no edges and $N$ as its set of vertices. Both $\mathsf{Set}$ and $\mathsf{Graph}$ are cocartesian monoidal and the functor $L \colon \mathsf{Set} \to \mathsf{Graph}$ is left adjoint to the forgetful functor $R \colon \mathsf{Graph} \to \mathsf{Set}$ which assigns to a graph $G$ its underlying set of vertices $U(G)$. Using structured cospans and appealing to Theorem \ref{SC}, we get a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{Graph})$ which has:
\begin{enumerate}
\item{sets as objects,}
\item{functions as vertical 1-morphisms,}
\item{cospans of graphs, or, \emph{open} graphs of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(N)$};
\node (B) at (1,0) {$G$};
\node (C) at (2,0) {$L(M)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I$} (B)
(C) edge node[above]{$O$} (B);
\end{tikzpicture}
\]
as horizontal 1-cells, where $L(N)$ and $L(M)$ are discrete graphs on the sets $N$ and $M$, respectively, $G$ is a graph and $I$ and $O$ are graph morphisms, and}
\item{maps of cospans of graphs of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(N_1)$};
\node (B) at (1,0) {$G_1$};
\node (C) at (2,0) {$L(M_1)$};
\node (A') at (0,-1) {$L(N_2)$};
\node (B') at (1,-1) {$G_2$};
\node (C') at (2,-1) {$L(M_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I_1$} (B)
(C) edge node[above]{$O_1$} (B)
(A') edge node[above]{$I_2$} (B')
(C') edge node[above]{$O_2$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [left]{$L(g)$} (C');
\end{tikzpicture}
\]
as 2-morphisms, where $L(f)$ and $L(g)$ are maps of discrete graphs induced by the underlying functions $f$ and $g$, respectively, and $\alpha \colon G_1 \to G_2$ is a graph morphism.
}
\end{enumerate}

>>>>

\begin{comment}
We can also define the part of the double equivalence $\mathbb{G} \colon F\mathbb{C}\textnormal{sp} \to {_L \mathbb{C} \textnormal{sp}(\mathsf{X})}$ which goes in the other direction: again, the object component of this double functor will be $\mathbb{G}_0 = \id_{\mathsf{A}}$.

@@

On the other hand, given a horizontal 1-cell $N$ in $F\mathbb{C}\textnormal{sp}$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$c_2$};
\node (D) at (1,-0.5) {$x \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I$} (B)
(C) edge node[above]{$O$} (B);
\end{tikzpicture}
\]
the horizontal 1-cell $\mathbb{G}(N)$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(c_1)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(c_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$!_c L(I)$} (B)
(C) edge node[above]{$!_c L(O)$} (B);
\end{tikzpicture}
\]
and then the horizontal 1-cell $\mathbb{E} \mathbb{G}(N)$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1.5,0) {$R(x)$};
\node (C) at (3,0) {$c_2$};
\node (D) at (1.5,-0.5) {$x \in F(R(x))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$R(!_c L(I))\eta_{c_1}$} (B)
(C) edge node[above]{$R(!_c L(O))\eta_{c_2}$} (B);
\end{tikzpicture}
\]
Then we have a 2-isomorphism $\epsilon_N \colon \mathbb{E} \mathbb{G} (N) \xrightarrow{\sim} N$ in $F\mathbb{C}\textnormal{sp}$ given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (D) at (1.5,0.5) {$x \in F(R(x))$};
\node (D') at (1.5,-1.5) {$x \in F(c)$};
\node (A) at (0,0) {$c_1$};
\node (B) at (1.5,0) {$R(x)$};
\node (C) at (3,0) {$c_2$};
\node (A') at (0,-1) {$c_1$};
\node (B') at (1.5,-1) {$c$};
\node (C') at (3,-1) {$c_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$R(!_c L(I))\eta_{c_1}$} (B)
(C) edge node[above]{$R(!_c L(O))\eta_{c_2}$} (B)
(A') edge node[above]{$I$} (B')
(C') edge node[above]{$O$} (B')
(A) edge node [left]{$1$} (A')
(B) edge node [left]{$e$} (B')
(C) edge node [left]{$1$} (C');
\end{tikzpicture}
\]
$$\iota \colon F(e)(x) \to x$$
where 
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1,0) {$R(L(c_1))$};
\node (C) at (2.5,0) {$R(L(c))$};
\node (D) at (3.5,0) {$R(x)$};
\node (E) at (4.5,0) {$c$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$\eta_{c_1}$} (B)
(B) edge node[above]{$R(L(I))$} (C)
(C) edge node[above]{$R(!_c)$} (D)
(D) edge node[above]{$e$} (E)
(A) edge[bend right] node [above]{$I$} (E);
\end{tikzpicture}
\]
commutes.
\end{comment}

\begin{comment}
\section{Applications}\label{Applications}
In this section we present several examples each of which may be realized in the context of decorated cospans or in the context of structured cospans. The first example regarding graphs was mentioned in the introduction and used as a reoccurring theme throughout the paper. The next three examples which take on more of an applied flavor, consists of electrical circuits, Markov processes and Petri nets. Each of these has been studied extensively by Baez, Fong, Master and Pollard by way of `black-boxing' \cite{BCR,BF,BFP,BM,BP}. Black-boxing is a way of interpreting the behavior of an open system, that is, a system with prescribed inputs and outputs such as the terminals of an electrical circuit, by observing the activity at the inputs and the outputs. The semantics of the activity at an open system's inputs and outputs is typically described in a category such as $\mathsf{LinRel}$ of finite dimensonal vector spaces and linear relations. Thus, in each case, black-boxing results in functors such as: $$\blacksquare_1 \colon \mathsf{Circ} \to \mathsf{LinRel}$$ $$\blacksquare_2 \colon \mathsf{Mark} \to \mathsf{LinRel}$$ $$\blacksquare_3 \colon \mathsf{Petri} \to \mathsf{LinRel}.$$ Each of these black-boxing functors also possess other convenient properties such as being symmetric monoidal. The first two of these were first done using Fong's theory of decorated cospans and then extended by the first two authors using structured cospans. The last two of these were also extended by being realized as double functors between double categories in recent works \cite{BC,BM}.
\subsection{Graphs}
As a first example that was also mentioned in the introduction, let $L \colon \mathsf{Set} \to \mathsf{Graph}$ be the functor that assigns to a set $N$ the \emph{discrete graph} on $N$ which is the edgeless graph $L(N)$ with no edges and $N$ as its set of vertices. Both $\mathsf{Set}$ and $\mathsf{Graph}$ are cocartesian monoidal and the functor $L \colon \mathsf{Set} \to \mathsf{Graph}$ is left adjoint to the forgetful functor $R \colon \mathsf{Graph} \to \mathsf{Set}$ which assigns to a graph $G$ its underlying set of vertices $U(G)$. Using structured cospans and appealing to Theorem \ref{SC}, we get a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{Graph})$ which has:
\begin{enumerate}
\item{sets as objects,}
\item{functions as vertical 1-morphisms,}
\item{cospans of graphs, or, \emph{open} graphs of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(N)$};
\node (B) at (1,0) {$G$};
\node (C) at (2,0) {$L(M)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I$} (B)
(C) edge node[above]{$O$} (B);
\end{tikzpicture}
\]
as horizontal 1-cells, where $L(N)$ and $L(M)$ are discrete graphs on the sets $N$ and $M$, respectively, $G$ is a graph and $I$ and $O$ are graph morphisms, and}
\item{maps of cospans of graphs of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(N_1)$};
\node (B) at (1,0) {$G_1$};
\node (C) at (2,0) {$L(M_1)$};
\node (A') at (0,-1) {$L(N_2)$};
\node (B') at (1,-1) {$G_2$};
\node (C') at (2,-1) {$L(M_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I_1$} (B)
(C) edge node[above]{$O_1$} (B)
(A') edge node[above]{$I_2$} (B')
(C') edge node[above]{$O_2$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [left]{$L(g)$} (C');
\end{tikzpicture}
\]
as 2-morphisms, where $L(f)$ and $L(g)$ are maps of discrete graphs induced by the underlying functions $f$ and $g$, respectively, and $\alpha \colon G_1 \to G_2$ is a graph morphism.
}
\end{enumerate}
